Synthesizing Symmetric Lenses

### 2019-01-31 papers/SymmetricLenses/main.tex
More polishing -- done for now

\newif\ifdraft\drafttrue  % set true to show comments
% \newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\ifappendices\appendicesfalse

@@

\subsection{Stochastic Regular Expression Equivalences}
The $\Expand$ algorithm needs to be able to enumerate SREs that are
``equivalent'' to a given one.
However, existing work does not provide a means of reasoning about stochastic
regular expression equivalence\bcp{Does not provide a means of {\em
    reasoning} about it, or does not even define it?}. We extend the
\emph{star-semiring}~\cite{Droste2009} equivalences\bcp{again, we seem to be
assuming that all readers will already know what they are...} to stochastic regular
expressions (Figure~\ref{fig:ssr-sre}).

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  The proofs of this theorem, and subsequent theorems in this paper, appears in
  the appendix presented in the full version of the paper accompanying this
  submission.
\end{proof}

We prove these star-semiring equivalences sound, as those are the equivalences
we either traverse (with \Expand), or normalize across (as part of
\GreedySynth).

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is the expected number of bits required to
describe a random element in $S$ (formally $-\Sigma_{s\in S}P(s)\Log_2P(s)$). The
entropy of a stochastic regular expression can be computed directly from its
syntax, when the stochastic regular expression is unambiguous and contains no
empty subcomponents.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 0\\

@@

    $\EntropyOf{\PRegexOr{\Regex}{\RegexAlt}{\Probability}}$
    & =
    & $\Probability(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
  \end{tabular}
\end{center}
We have proven that this syntactic computation corresponds to the entropy of the
stochastic regular expression when the stochastic regular expression is
unambiguous (parses strings uniquely).

\begin{theorem}
  \label{thm:correct_entropy}
  If $\Regex$ is unambiguous and does not contain $\emptyset$ as a subterm,
  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.
\end{theorem}

Without unambiguity constraints it is not easy to calculate entropy (and the
method we have given is certainly incorrect).  For example, consider the SRE
$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,
this SRE has an entropy of $1$, but the entropy should be $0$. These
difficulties become particularlly complex when the languages are ambiguous in
subtle ways and the algorithm must be correct when only some of the two languages
overlap. Similar issues occur when trying to find the entropy when $\emptyset$
is a subterm (what should the entropy of $\PRegexStar{\emptyset}{.5}$ be?),
so we do not define entropy on $\emptyset$.

Fortunately, we already require unambiguous regular expressions as input to
our synthesis procedure for other reasons, and we can easily preprocess
emptysets out of the stochastic regular expressions that are themselves
nonempty by traversing the appropriate star-semiring equivalences (\EG,
\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).

@@

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. While our algorithms perform synthesis on stochastic regular
expressions, lenses themselves are typed between regular expressions. We can
recover a regular expression from a stochastic regular expression by deleting
probability annotations. In a slight abuse of notation, if $\Regex$ is a
stochastic regular expression, $\BRegex$ is the regular expression obtained by
removing probability annotations. The typing judgment $\Lens \OfType \BRegex
\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.

@@

We have already described a number of these lenses in \S\ref{sec:overview}.
The compose lens $\Lens_1 ; \Lens_2$ composes two lenses sequentially, it
transforms from one data format to the other by first transforming to an
intermediate format, shared as the target of $\Lens_1$ and the source of
$\Lens_2$. The \MergeR lens combines two sublenses which share the same target.
Which lens gets used depends on what data is in the left-hand format, if it is
in the domain of $\Lens_1$ then $\Lens_1$ gets used, and similarly for
$\Lens_2$.  The \MergeL lens is symmetric to \MergeR, the two sublenses must
share the same source. Finally, $\InvertOf{\Lens}$ inverts a lens, \CreateR
becomes \CreateL and \PutR becomes \PutL, and vice-versa.

We now provide the typing rules and semantics for these combinators. We do
not include the semantics for combinators equivalent to those presented in
prior work~\cite{symmetric-lenses,Boomerang}. Furthermore, we only present
\CreateR and \PutR when \CreateL and \PutL are symmetric; full details can
be found in an appendix\ifappendices\else{} (in the supplemental
materials)\fi. For simplicity of presentation, we use unambiguous regular
expressions everywhere, even though some lenses can actually be defined over
ambiguous regular expressions (for example, the identity lens doesn't need
any ambiguity constraints\bcp{Is this the only example?}); we believe
relaxing this restriction is unproblematic.

@@

    $\PutR{} \App \String \App \StringAlt$
    & =
    & $\begin{cases*}
        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$
      \end{cases*}$\\
  \end{tabular}
\end{tabular}
\end{centermath}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
\begin{centermath}
  \begin{tabular}{m{5cm}m{6cm}}
$
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
  }
  {
    \MergeROf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\BRegex_1}{\BRegex_2}
    \Leftrightarrow
    \BRegexAlt
  }
$
&
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\

@@

\subsection{Lens Costs}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds of the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. We use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We use the notation $a(b,c)$ to
mean $(a*b,a*c)$, and the notation $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

@@

    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
    & =
    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
  \end{tabular}
\end{center}
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically. These
functions bound the expected number of bits to recover one data format from a
synchronized string in the other format. Note that we would be able to exactly
calculate the conditional entropy, were it not for \MergeL. If
$\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
need to determine if the synchronized string is in $\RegexAlt_1$ or $\RegexAlt_2$.
However, this information content is dependent on the how likely the
synchronized string is to be in $\RegexAlt_1$ or $\RegexAlt_2$. Nevertheless, we
typically calculate the conditional entropy exactly, as merges are relatively
uncommon in practice; only 2 of the lenses sythesized in our benchmarks
include merges.

@@

Note that our definition of $\Entropy^\rightarrow$ contains no case for sequential
composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
such compositions. Unfortunately, defining the entropy of lenses involving
composition is challenging because $\Lens_1$ might, for instance, add some
information that is subsequently projected away in $\Lens_2$. Such operations
can cancel, leaving a zero-entropy bijection composed from two non-zero entropy
transformations. However, detecting such cancellations directly is complicated and
this property is difficult to determine merely from syntax. Fortunately, we are able to
avoid such complication altogether by synthesizing \emph{DNF lenses}---simple symmetric
lenses that inhabit a disjunctive normal form. DNF lenses have a number of
properties that facilitate synthesis: they cut down the program search space
dramatically and they do not contain instances of composition.

@@

The details of these definitions are syntactically tedious, but not
intellectually difficult. We elide them here but include them in the appendix.

@@

\paragraph*{Relevance Annotations}
In this work, there are three competing constraints. The tool should (1) return
lenses quickly to users, (2) return the highest quality lenses, and (3)
not require users to write a large number of examples.
Our termination heuristic works well for most situations, but
there are cases where it fails.

@@

Conversely, the termination heuristic is satisfied algorithm cuts short: if it
had continued searching for lenses it would have found the desired one, but the
termination heuristic was satisfied ``too early.'' We wish for our algorithm to
provide feedback in a timely manner, but sometimes this prohibits finding the
desired lens.

@@

If synthesizing a lens seems to be taking a long time, it is spending time
checking it cannot find a lens that projects less information. By ``skipping''
portions of the SRE that \emph{should} be projected, the algorithm will not
continue searching for a better lens when it has already found the correct one.
In practice, we implement \Skip by setting the information content of
$\Skip(\Regex)$ to zero.

If the algorithm returns a lens with too disconnects and merges, one could
provide additional examples to constrain the lens more. However, for large
formats, writing such examples is time consuming and error-prone. Instead, if
one recognizes that information from $\Regex$ is incorrectly lost when converted
between formats, they can prevent that information from being lost by annotating
it $\SRequireOf{\Regex}$. One can guarantee a lens between \Regex and \RegexAlt
is bijective by annotating them as $\SRequireOf{\Regex}$ and
$\SRequireOf{\RegexAlt}$. In practice, we implement \SRequire by setting the
cost of any lens that loses ``required'' information to infinity.
% end synthesis

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented symmetric DNF lenses as an extension to
Boomerang. Specifically, we added symmetric lens combinators to
Boomerang, and implemented all of Boomerang's asymmetric lens
combinators  using a combination of the symmetric lens combinators presented in this
paper and symmetric versions of the asymmetric extensions (like matching lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses})
already present in Boomerang.
We also integrated our synthesis engine into Boomerang. This
integration allows users to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into 
manually-written lenses, and reference previously defined lenses in
the synthesis of new lenses.

@@

\paragraph*{Relation with Simple Symmetric Lenses} To compare classical and simple
symmetric lenses, we define an $apply$ function on simple symmetric lenses as well. If
$apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
$es_i$, the lens \Lens generates $es_i'$ (the right format if $es_i =
\InLOf{x}$, and the left format if $es_i = \InROf{y}$). If
$apply(\Lens,\SomeOf{(x,y)},es) = es'$, then starting with data $x$
and $y$ on the left and right, respectively, after edit $es_i$, the lens \Lens
generates $es_i'$.

@@

\paragraph*{Other Lens Formulations}
Bidirectional programming has an extensive literature, with many extensions onto
basic lens formulations, like quotient lenses\cite{quotientlenses} and matching
lenses\cite{matchinglenses}. Interested readers can consult
\citet{DBLP:conf/icmt/CzarneckiFHLST09} for a slightly dated survey of lenses
and other lens-like structures.

@@

Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
\begin{align*}
H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
\end{align*}
\item
($S = S_1 \; |_p \; S_2$)

>>>>

\newif\ifdraft\drafttrue  % set true to show comments
% \newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\ifappendices\appendicestrue

@@

\subsection{Stochastic Regular Expression Equivalences}
The $\Expand$ algorithm needs to be able to enumerate SREs that are
``equivalent'' to a given one.
However, existing work does not provide a means of reasoning about stochastic
regular expression equivalence\bcp{Does not provide a means of {\em
    reasoning} about it, or does not even define
  it?}. Figure~\ref{fig:ssr-sre} shows how we to extend the
\emph{star-semiring}~\cite{Droste2009} equivalences\bcp{again, we seem to be
assuming that all readers will already know what they are...} to stochastic regular
expressions.\bcp{Notation: I don't see any reason for the $s$ superscript in
$\SSREquiv$}

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  See the appendix%
\ifappendices\else{} included with the supplemental
material\bcp{Right?}\fi.  \bcp{It took me quite a while to find it.  We
  should give an actual pointer to a specific theorem in the appendix.}
\end{proof}

This theorem will come in handy as we traverse (with \Expand) or normalize
across (as part of \GreedySynth) star-semiring equivalences.

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is the expected number of bits required to
describe a random element in $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot \Log_2P(s)$). The
entropy of a stochastic regular expression can be computed directly from its
syntax, when it is unambiguous\bcp{undefined term} and contains no
empty subcomponents.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 0\\

@@

    $\EntropyOf{\PRegexOr{\Regex}{\RegexAlt}{\Probability}}$
    & =
    & $\Probability(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
  \end{tabular}
\end{center}
\bcp{This sentence says exactly what the previous one does.  Drop.}We have proven that this syntactic computation corresponds to the entropy of the
stochastic regular expression when the stochastic regular expression is
unambiguous (parses strings uniquely).

\bcp{Again, this is a critical definition for understanding the paper, and
  it is pretty gnarly.  Examples, please!}

\begin{theorem}
  \label{thm:correct_entropy}
  If $\Regex$ is unambiguous and does not contain $\emptyset$ as a subterm,
  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.\bcp{Should it not be $-\EntropyOf{\Regex}$??}
\end{theorem}

Without unambiguity constraints, the method above for calculating entropy does
not work, and it is not easy to see how to generalize it.  For example,
consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using
the above computation, this SRE has an entropy of $1$, but the entropy
should be $0$. These difficulties become particularly complex when the
languages are ambiguous in subtle ways and the algorithm must be correct
when only some of the two languages overlap. Similar issues occur when
trying to find the entropy when $\emptyset$ is a subterm (what should the
entropy of $\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy
on~$\emptyset$.

Fortunately, we already require unambiguous regular expressions as input to
our synthesis procedure for other reasons, and we can easily preprocess
empty subexpressions out of SREs that are themselves nonempty using the
star-semiring equivalences (\EG, $\PRegexOr{\emptyset}{\lstinline{"s"}}{.5}
\SSREquiv \lstinline{"s"}$).

@@

%%%%%%%%%%%%%%  BCP: stopped reading carefully here...

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. While our algorithms perform synthesis on stochastic regular
expressions, lenses themselves are typed between regular expressions. We can
recover a regular expression from a stochastic regular expression by deleting
probability annotations. In a slight abuse of notation\bcp{why is this an
  abuse of notation??}, if $\Regex$ is a
stochastic regular expression, $\BRegex$ is the regular expression obtained by
removing probability annotations. The typing judgment $\Lens \OfType \BRegex
\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.  \bcp{Ugly typesetting:}

@@

We have already described a number of these combinators in \S\ref{sec:overview}.
Briefly (full definitions are below), the compose lens $\Lens_1 ; \Lens_2$
composes two lenses sequentially, it 
transforms from one data format to the other by first transforming to an
intermediate format, shared as the target of $\Lens_1$ and the source of
$\Lens_2$. The \MergeR lens combines two sublenses which share the same target.
Which lens gets used depends on what data is in the left-hand format, if it is
in the domain of $\Lens_1$ then $\Lens_1$ gets used, and similarly for
$\Lens_2$.  The \MergeL lens is symmetric to \MergeR, the two sublenses must
share the same source. Finally, $\InvertOf{\Lens}$ inverts a lens, \CreateR
becomes \CreateL and \PutR becomes \PutL, and vice-versa.

We now give typing rules and semantics for these combinators. We do not
include the semantics for combinators equivalent to those presented in prior
work~\cite{symmetric-lenses,Boomerang}\bcp{Huh?}. Furthermore, we present
just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
full details can be found in an appendix\bcp{where,
  exactly??}\ifappendices\else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions
everywhere, even though some lenses can actually be defined over ambiguous
regular expressions (for example, the identity lens doesn't need any
ambiguity constraints\bcp{Is this the only example?}); we believe relaxing
this restriction is unproblematic.

@@

    $\PutR{} \App \String \App \StringAlt$
    & =
    & $\begin{cases*}
        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$
      \end{cases*}$\\
  \end{tabular}
\end{tabular}
\end{centermath}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.  \bcp{We should highlight here the fact that this is a significant
  loss of power for some examples, compared to full symmetric lenses.}
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
\begin{centermath}
  \begin{tabular}{m{5cm}m{6cm}}
$
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
  }
  {
    \MergeROf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\BRegex_1}{\BRegex_2}
    \Leftrightarrow
    \BRegexAlt
  }
$
&
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\

@@

\subsection{Lens Costs}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds of the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

@@

    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
    & =
    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
  \end{tabular}
\end{center}
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically. These
functions bound the expected number of bits to recover one data format from a
synchronized string in the other format. Note that we would be able to exactly
calculate the conditional entropy, were it not for \MergeL\bcp{and \MergeR??}. If
$\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
need to determine if the synchronized string is in $\RegexAlt_1$ or $\RegexAlt_2$.
However, this information content is dependent on the how likely the
synchronized string is to be in $\RegexAlt_1$ or $\RegexAlt_2$. Nevertheless, we
typically calculate the conditional entropy exactly, as merges are relatively
uncommon in practice; only 2 of the lenses sythesized in our benchmarks
include merges.

@@

Note that our definition of $\Entropy^\rightarrow$ contains no case for sequential
composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
such compositions.  Defining the entropy of lenses involving
composition is challenging because $\Lens_1$ might, for instance, add some
information that is subsequently projected away in $\Lens_2$. Such operations
can cancel, leaving a zero-entropy bijection composed from two non-zero entropy
transformations. However, detecting such cancellations directly is complicated and
this property is difficult to determine merely from syntax. Fortunately, we are able to
sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
lenses that inhabit a disjunctive normal form. DNF lenses have a number of
properties that facilitate synthesis: they cut down the program search space
dramatically and they do not contain instances of composition.\bcp{The last
  sentence basically repeats what the rest of the paragraph already says.}

@@

The details of these definitions are syntactically tedious, but not
intellectually difficult. We elide them here but include them in the appendix.\bcp{Where?}

@@

\paragraph*{Relevance Annotations}\bcp{Wonder whether this should be its own
top-level section...?}
In this work, there are three competing constraints. The tool should (1) return
lenses quickly to users, (2) return the highest quality lenses, and (3)
not require users to write a large number of examples.
Our termination heuristic works well for most situations, but
there are cases where it fails.

@@

Conversely, the termination heuristic is satisfied algorithm cuts
short\bcp{huh??}: if it 
had continued searching for lenses it would have found the desired one, but the
termination heuristic was satisfied ``too early.'' We wish for our algorithm to
provide feedback in a timely manner, but sometimes this prohibits finding the
desired lens.

@@

If synthesizing a lens seems to be taking a long time, it is\bcp{always?} spending time
checking it cannot find a lens that projects less information. By ``skipping''
portions of the SRE that \emph{should} be projected, the algorithm will not
continue searching for a better lens when it has already found the correct one.
In practice, we implement \Skip by setting the information content of
$\Skip(\Regex)$ to zero.

If the algorithm returns a lens with too disconnects\bcp{??} and merges, one could
provide additional examples to constrain the lens more. However, for large
formats, writing such examples is time consuming and error-prone. Instead, if
one recognizes that information from $\Regex$ is incorrectly lost when converted
between formats, they can prevent that information from being lost by annotating
it $\SRequireOf{\Regex}$. One can guarantee a lens between \Regex and \RegexAlt
is bijective by annotating them as $\SRequireOf{\Regex}$ and
$\SRequireOf{\RegexAlt}$. In practice, we implement \SRequire by setting the
cost of any lens that loses ``required'' information to infinity.

\bcp{This discussion seems both too brief---maybe some concrete examples
  would help?---and too informal: E..g., What do skip and require have to do
  with probability distributions and entropy?}

% end synthesis

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented symmetric DNF lenses as an extension to
Boomerang. Specifically, we added symmetric lens combinators to
Boomerang\bcp{that's exactly what the previous sentence said!}, and
implemented all of Boomerang's asymmetric lens 
combinators  using a combination of the symmetric lens combinators presented in this
paper and symmetric versions of the asymmetric extensions (like matching lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses})
already present in Boomerang.
We also integrated our synthesis engine into Boomerang. This
integration allows users to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into 
manually-written lenses, and reference previously defined lenses in
the synthesis of new lenses.

@@

\paragraph*{Symmetric Lenses vs.{} Simple Symmetric Lenses} To compare classical and simple
symmetric lenses, we define an $apply$ function on simple symmetric lenses as well. If
$apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
$es_i$, the lens \Lens generates $es_i'$ (the right format if $es_i =
\InLOf{x}$, and the left format if $es_i = \InROf{y}$). If
$apply(\Lens,\SomeOf{(x,y)},es) = es'$, then starting with data $x$
and $y$ on the left and right, respectively, after edit $es_i$, the lens \Lens
generates $es_i'$.

@@

\paragraph*{Other Lens Formulations}\bcp{This paragraph doesn't add much}
Bidirectional programming has an extensive literature, with many extensions onto
basic lens formulations, like quotient lenses\cite{quotientlenses} and matching
lenses\cite{matchinglenses}. Interested readers can consult
\citet{DBLP:conf/icmt/CzarneckiFHLST09} for a slightly dated survey of lenses
and other lens-like structures.

@@

Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
  signs in the last few lines here.}
\begin{align*}
H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
\end{align*}
\item
($S = S_1 \; |_p \; S_2$)
